Kelp morphometric properties and its link to the hydrodynamic modelling

### 2018-08-01 chapter_1.Rmd
removed some sections. I am keeping the background literature with regards to biotic drivers and overall kelp ecology for now. I think it may be important for the reader to understand the complexity of temperate reefs.

---
title: "Kelp morphometric properties and its link to the hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  word_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
bibliography: mycollection2.bib
---

@@

Light is an important factor for kelp survival, however if light is limited or excessive this may negatively impact kelp survival or growth. Much of the past research into the role light plays into the functioning of kelp [@Bruhn1996);@Belseth2012]. For instance, solar ultraviolate radiation has been shown to aﬀect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood 1987]. The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reducedgrowthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987]. Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the eﬀects mentioned above. High light stress has negative eﬀects, such as photoinhibition and photo-damage on *Ecklonia cava* sporophytes [@Altamirano2004]. @Altamirano2004 found that *Ecklonia cava* is more vulnerable to light stress conditions, and less likely to recover under unfavourable conditions [@Altamirano2004]. @Bolton1985 showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production. A interesting ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons. The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985]. Light is also important for survival and functioning of benthic algal species; a study by @Toohey2004 investiagted the possible link of light and thallus scour of the kelp *Ecklonia radiata canopy* on the associated foliose algal assemblage. The authors concluded that a reduction in light, rather than thallus scour has an eﬀect on foliose algal assemblage. Reduced light was associated with reduced diveristy and an marked increase abundance in certian foliose species, which were able to acclimiate more light under low irradiances compared to other species [@Toohey2004]. Kelp may also compete intra-specﬁcally or inter-speciﬁcally for light. For instance, @Clark2004 investiagted eﬀect of canopy shading of *Macrocystis pyrifera* and understorey canopy shading of *Pterygophora californica* individually and in combination of benthic macro-algae assemblage. Under all treatments there was a sigiﬁance change in macro-algal assemblage, with certian species able to recover in abundance more readily than others. The authors devided these species into two categories; “shade-tolerant” and “light adapted”, and conlcuded that certian foliose species are better adapted for diﬀerent light conidtions in accordance with canopy cover. In South Africa, the eﬀect of *Ecklonia maxima* harvesting on sub-story kelp, kelp population structure and kelp recruitement was investigated by @Rothman2006. The results showed no signiﬁcant diﬀerence in the three factors mentioned earlier and @Rothman2006 concluded that factors other than light play a role in structuring kelp communities.

@@

Nutrients Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood [@Dayton1985; @Gaylord2012]. Dissolved nitrogen, and in particular nitrate, are important; however research has also placed emphasis on phosphate and other trace compounds for functioning of kelps [@Dayton1985]. Additionally, some kelps have the ability to store inorganic nitrogen in order to compensate for periods of low nutrient availability, which has been observed for Laminaria and Macrocystis [@Dayton1985; @Gaylord2012]. Nutrient stratiﬁcation is also an important factor, particularly for canopy type kelps. The concentration of nutrients at the surface is important to the functioning and maintenance of the canopy. For instance kelp canopies in California often deteriorate in the summer months when surface nitrate levels are low [@Jackson1977]. Water motion is important in the assimilation of nutrients from the water column, and kelps have been shown to adapt blade morphology in order to create more turbulence around the boundary layer of the frond to enhance nutrient assimilation (Wheeler 1980). Temperature has also been closely linked with nutrient concentrations. Nutrients are often in higher concentrations in the water column during low temperature events. This is often an indication of an “up-welling” event, which brings cold and nutrient rich waters from the bottom to the surface of the water column. However temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988]. However most studies investigate multiple characteristics in one study (i.e. temperature and light and water motion), while studies investigating the direct role of temperature in nutrient assimilation is lacking in marine algae is lacking.

@@

#Sedimentation

Sedimentation and scour can be highly destructive to kelp forests with most detrimental impacts on the development stages such as spores, gametophytes and young plants (Dayton 1985), and has been linked to seasonal trends relating to wind direction [@Moore1972]. @Devinny1978 found that sedimentation greatly reduced spore settlement and which in turn greatly reduces the probability of survival for Macrocystis Pyrifera gametophytes. Some species seem more resistant to the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea compared to others such as Macrocystis Pyrifera (Dayton 1985). The impacts of sedimentation is not limited to the sub-tidal but has also been shown to eﬀect inter-tidal kelp as well. For example a study by [@Schiel2006] investigated the eﬀects of sedimentation on the post-settlement stages of habitat-forming seaweeds. The results showed high mortality of germlings at sites that had high sedimentation, and was further supported through laboratory tests. Grazing was found to be negligible and mortality of gametophytes was attributed to sedimentation, although its eﬀects did vary with molluscan grazing [@Schiel2006]. The presence of kelps themselves can also eﬀect the degree of sedimentation. For instance, @Wernberg2005 investigated the ability of Ecklonia radiata (canopy forming species) to modify the physical environment and the related eﬀects on associated foliose algae. An increase in kelp densities was marked by a signiﬁcant reduction of light and sedimentation, and found that sedimentation did not have a signiﬁcant eﬀect on associated foliose algae [@Wernberg2005]. Therefore it is often diﬃcult to separate the direct eﬀects of each of the environmental variables as they are often related to one another. However, sea surface temperatures is one of the main predictors of marine biological systems, as it has been shown to relate to physiological and ecological processes at all levels [@Andersen2013; @Bolton1985; @Bolton2012; @Dayton1985]. Kelp forests are no exception, with temperature being the main limiting factor [@Dayton 1985].

@@

>>>>

---
title: "Kelp morphometric properties and its link to the hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  word_document:
    toc: yes
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: yes
---

@@

Light is an important factor for kelp survival, however if light is limited or excessive this may negatively impact kelp survival or growth. Much of the past research into the role light plays into the functioning of kelp [@Bruhn1996);@Belseth2012]. For instance, solar ultraviolate radiation has been shown to aﬀect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood 1987]. The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reduced growthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987]. Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the eﬀects mentioned above. High light stress has negative eﬀects, such as photoinhibition and photo-damage on *Ecklonia cava* sporophytes [@Altamirano2004]. @Altamirano2004 found that *Ecklonia cava* is more vulnerable to light stress conditions, and less likely to recover under unfavourable conditions [@Altamirano2004]. @Bolton1985 showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production. A interesting ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons. The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985].

@@

The importance of nutrients in the functioning of kelps is well understood [@Dayton1985; @Gaylord2012]. Dissolved nitrogen, and in particular nitrate, are important; however research has also placed emphasis on phosphate and other trace compounds for functioning of kelps [@Dayton1985]. Additionally, some kelps have the ability to store inorganic nitrogen in order to compensate for periods of low nutrient availability, which has been observed for Laminaria and Macrocystis [@Dayton1985; @Gaylord2012]. Nutrient stratiﬁcation is also an important factor, particularly for canopy type kelps. The concentration of nutrients at the surface is important to the functioning and maintenance of the canopy. For instance kelp canopies in California often deteriorate in the summer months when surface nitrate levels are low [@Jackson1977]. Water motion is important in the assimilation of nutrients from the water column, and kelps have been shown to adapt blade morphology in order to create more turbulence around the boundary layer of the frond to enhance nutrient assimilation (Wheeler 1980). Temperature has also been closely linked with nutrient concentrations. Nutrients are often in higher concentrations in the water column during low temperature events. This is often an indication of an “up-welling” event, which brings cold and nutrient rich waters from the bottom to the surface of the water column. However temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988]. However most studies investigate multiple characteristics in one study (i.e. temperature and light and water motion), while studies investigating the direct role of temperature in nutrient assimilation is lacking in marine algae is lacking.

@@

@@

# Wave exposure
